Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2015-05-24 unstructured.tex
2015.05.18 (Alex) Gone through spelling correction. Added minimal biography text for authors. Some more re-wording on the connection sentences in the discussion section. Moving table and plots to improve overall typesetting in the structured section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \paragraph*{ \bf Delay Aware P2P System}
A \emph{delay--aware} \p\ system termed {\sl DAPS} 
was introduced in~\cite{ZL2005}.
{\sl DAPS}  seeks to attain reduced look-up times by dividing 
peer routing tables into several sectors of increasing delay. 
The source node that issues the query designates the delay 
boundary it may tolerate, providing so a \emph{pruning factor}.
In this context, user requests are forwarded only to 
nodes whose expected delay is less than or equal to the indicated boundary. 
{\sl DAPS} primarily focuses on user ``experience'' and deploys an
end-to-end delay monitoring mechanism that may enable
the clustering of routing tables. 
In an dynamic environment, the formation 
of very accurate such routing tables might be however an elusive goal.
%With the clustered
%routing tables and the loose organization of DAPS' overlay network it is
%considered by is be tween structured and unstructured.
%%
In terms of the $3$ performance criteria, {\sl DAPS} stands as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Does not check for close by nodes, it just sorts the contents of a node's
% routing table.
% Also in a high churn environment a lot of time could be spend probing for
% distances, resorting routing tables and creating sectors. Moreover it just
% prunes distant neighbout
low &
% Sorting routing table entries is done locally so it is not affecting.
low &
%
medium
\end{tabular}
}
\end{center}

@@

Along with {\sl Narada}, the \emph{Adaptive Overlay Topology 
Optimization (AOTO)} \cite{LZXN2003} is one of the first attempts to address the
topology mismatch problem. \emph{AOTO} is a distributed algorithm that seeks to
optimize the usage of the underlying physical resources and operates in
$2$ phases:
\emph{Selective Flooding} and \emph{Active Topology}.
In \emph{Selective Flooding}, a minimum spanning tree is built for each peer and
its immediate neighbors so that queries do not flood the entire network and at
the same time preventing their search scope from shrinking. This way some
neighbors become non--flooding. During \emph{Active Topology}, each peer
replaces independently such non--flooding neighbors, with closer nodes as an attempt
to revise the overlay links so that they can reflect more closely the physical
network topology. Picking a replacement out of these non-flooding
neighbors follows a random policy called \emph{Randomized AT} algorithm.
To accomplish the above actions, a peer has to keep track of 
its communication costs with all its neighbors (e.g., network delays)
as well as the costs between any pair of neighbors. The \emph{randomized AT}
algorithm is applied by a source peer
every time its neighbor list is updated or an updated neighbor cost 
table is received.
%%
%\paragraph*{Selective Flooding (SF)}
%
% TODO: I DON'T REMEMBER WHY THE FOLLOWING IS MENTIONING LTM WHICH IS ANOTHER
%       ALGORITM. MAYBE THIS CAN BE USED AS A PART OF DISCUSSION AND/OR
%       COMPARISON OF AOTO AND LTM
% TODO: EDIT: LTM SEEMS A MISTAKE HERE BETTER NOT USE IT FOR DISCUSSION (THIS
%             IS AOTO)
%LTM's SF effectiveness has be proven to be detached from the different physical
% or overlay topologies. On the other hand, SF is more effective with large
%number of logical neighbors. It can reach an average optimization rate of 87.4
%percent on a logical topology with an average of 30 logical neighbors.
%
%\paragraph*{Active Topology (AT)}
%
%Different numbers of average logical neighbors has little to do with the
% effectiveness of AT. If the source has $n$ non-flooding peers, there are $n$
%potential neighbor replacements. The overhead to exhaust all $n$ possible
%replacements can be high, so in practice, after each replacement the source
%peer can decide whether it needs to find another candidate peer. This is done
%by computing the cost improvement ratio greater than some predefined
%termination threshold. The larger the threshold, the slower, in the number of
%optimization steps, the reduction of the normalized average distance. As a
%whole the average response time is significantly reduced when more optimization
%steps taken.
%%
\emph{AOTO}'s performance regarding the $3$ criteria is as follows:
%%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
medium &
% tracking of communication costs to all peer's neighbors as well as between
% any pair of neighbors. Whenever a new neighbor cost table is received or
% there is a change of neighbors, the source peer has to re-calculate the
% multicast tree and apply the randomized AT algorithm.
high &
% needs global knowledge
medium
\end{tabular}
}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The main objective of the \emph{Biased Neighbor Selection (BNS)} 
approach~\cite{BCCMSBZ2006} is to 
strengthen {\sl BitTorrent}'s~\cite{c_bittorrent_2003} locality 
by carefully choosing most of the neighbors of a peer to come from 
the same \emph{ISP}, while simultaneously adhering to the near-optimal
download performance of the protocol.
%%
By and large, {\sl BitTorrent} deploys mechanisms that have proved
aggressive to \emph{ISP} networking and accounting as they function 
in a ``without-borders'' approach. 
The default {\sl BitTorrent} specification
designates $35$ connections for each peer regardless if such connections are
placed within or outside the borders of the \emph{ISP} that hosts the node.
%%
\emph{BNS} proposes that only $k$ out of these $35$ connections be
established with nodes beyond the {\it ISP}-borders;
the $k$ connections offer an extended and ``global'' view of 
the network at large and seek to strike a balance between 
the load imposed inside and outside the \emph{ISP}.
The latter refrains peers from exchanging traffic through
expensive network links should there exist alternative local connections
offering the required services. 
%%
\emph{BNS} is realized through 
\begin{inparaenum}[1)]
\item
the modification of the tracker so that 
\emph{ISP}-local resources are rapidly identified
\item
the use of \p--traffic--shaping devices on \emph{ISP} edge--routers. 
\end{inparaenum}
%%
In~\cite{BCCMSBZ2006}, a combination of 
biased neighbor selection along with bandwidth throttling
and caching techniques for near-optimal lookup results is proposed. 
The effectiveness of \emph{BNS} is debated 
in~\cite{RTLCGZ2010} where experimental results suggest that 
most peers in a swarm reside in different authonomous systems (\emph{AS}s).
Consequently, clients frequently do not have enough peers to leverage
intra-\emph{AS} connections \cite{RTLCGZ2010}.
Moreover, \emph{BNS} deployment
requires knowledge of \emph{ISP} network mappings and awareness of possible
changes that occur in the infrastructure limiting its widespread adoption; the
following table qualitatively depicts the expected behavior of \emph{BNS}:
%%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% inter-ISP cost minimized while retaining the near optimal performance of
% BitTorrent download.
high &
% this info is taken at swarm join
low &
% trackers should maintain and update proximity info/proximity info is
% retrieved from special hardware.
medium
\end{tabular}
}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Locality-Awareness in BitTorrent-like P2P Applications}
In~\cite{LCLX2009}, three techniques that 
``inject'' locality awareness in {\sl BitTorrent}-like applications are discussed.
The first \emph{macroscopic-level} technique
focuses on improving the neighbor selection process. 
When a peer asks a tracker to join, 
the later sorts its swarm peers according to
their hop-count distance from the requesting peer 
send out the top-$k$ list of nodes (e.g., $k$=$50$).
%%
The second technique, applied in an \emph{intermediate-level}, manipulates
{\sl BitTorrent}'s chocking/unchocking mechanism. A peer unchockes its $4$
closest, in terms of 
autonomous system hop-count, interested neighbors. 
The same applies also when the peer turns to 
the seeding state;  this  \emph{intermediate-level} approach favors
least distance and is in contrast to the original {\sl BitTorrent} 
implementation which favors uploading speed.
%%
The last technique operates at a \emph{microscopic-level} by chosing to pick
the next chunk to download based on a locality-first policy; this picks first
the chunk that is closer, as opposed to {\sl BitTorrent}'s rare-first chunk
picking policy.
%%VM I once more reworded the above, seemed to complex to me
The distance value of a chunk is computed as the mean
value of the distances of the peers that posses it.
Through experimentation on a real-world Internet topology simulated in
{\sf PlanetLab}, the locality-based approach achieves traffic optimization.
Nevertheless, it does not do that well when compared with the
random approach of the standard {\sl BitTorrent} protocol.

@@

Below, we outline how the proposed protocols use elements of the above four
methodologies. Then we offer a summary qualitative comparison for all surveyed
approaches applicable for \emph{unstructured}~\p~\emph{systems}.
Specifically, Table~\ref{unstructured:table} summarises, for each
effort, the methodologies it incorporates, its special highlights as well as 
a rough estimation of the pros and cons in its implementation. Last but
not least, with Figures~\ref{unstructured:plot:efficiency}, \ref{unstructured:plot:overhead}
and \ref{unstructured:plot:scalability} we pictorially compare the
algorithms in terms of \emph{efficiency}, \emph{overhead} and \emph{scalability}
respectively; criteria we defined in Subsection~\ref{background:motivation}
and which we used to evaluate the effectiveness of the each surveyed algorithm.

@@

In landmark-based algorithms, nodes use network delay (e.g., \emph{RTT}) as a
distance measurement method to position themselves with respect to ``a priori''
known servers on the Internet, like \emph{Ono} which uses the \emph{CDN}-infrastructure
for this purpose.
For example, in Figure~\ref{figure:landmarking}, the newly
arriving peer, measures its distance to an array of landmark points denoted as
$L_n$. Then it creates an ordered list of the landmarks based on its measured
distances uses that information to select its neighbours. Specifically,
the landmark--servers are used by nodes to estimate 
their positions based on the intuitive assumption that peers
with similar distances to a set of landmarks, are physically close to each
other, as well, over the network.  
%%
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.34]{img/pdf/landmarking.pdf}
\caption{Landmark binning during node bootstrap.}
\label{figure:landmarking}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \paragraph*{ \bf Delay Aware P2P System}
A \emph{delay--aware} \p\ system termed {\sl DAPS} 
was introduced in~\cite{ZL2005}.
{\sl DAPS}  seeks to attain reduced look-up times by dividing 
peer routing tables into several sectors of increasing delay. 
The source node that issues the query designates the delay 
boundary it may tolerate, providing so a \emph{pruning factor}.
In this context, user requests are forwarded only to 
nodes whose expected delay is less than or equal to the indicated boundary. 
{\sl DAPS} primarily focuses on user ``experience'' and deploys an
end-to-end delay monitoring mechanism that may enable
the clustering of routing tables. 
In an dynamic environment, the formation 
of very accurate such routing tables might be however an elusive goal.
%With the clustered
%routing tables and the loose organization of DAPS' overlay network it is
%considered by is be tween structured and unstructured.
%%
In terms of the $3$ performance criteria, {\sl DAPS} stands as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Does not check for close by nodes, it just sorts the contents of a node's
% routing table.
% Also in a high churn environment a lot of time could be spend probing for
% distances, resorting routing tables and creating sectors. Moreover it just
% prunes distant neighbor
low &
% Sorting routing table entries is done locally so it is not affecting.
low &
%
medium
\end{tabular}
}
\end{center}

@@

Along with {\sl Narada}, the \emph{Adaptive Overlay Topology 
Optimization (AOTO)} \cite{LZXN2003} is one of the first attempts to address the
topology mismatch problem. \emph{AOTO} is a distributed algorithm that seeks to
optimize the usage of the underlying physical resources and operates in
$2$ phases:
\emph{Selective Flooding} and \emph{Active Topology}.
In \emph{Selective Flooding}, a minimum spanning tree is built for each peer and
its immediate neighbors so that queries do not flood the entire network and at
the same time preventing their search scope from shrinking. This way some
neighbors become non--flooding. During \emph{Active Topology}, each peer
replaces independently such non--flooding neighbors, with closer nodes as an attempt
to revise the overlay links so that they can reflect more closely the physical
network topology. Picking a replacement out of these non-flooding
neighbors follows a random policy called \emph{Randomized AT} algorithm.
To accomplish the above actions, a peer has to keep track of 
its communication costs with all its neighbors (e.g., network delays)
as well as the costs between any pair of neighbors. The \emph{randomized AT}
algorithm is applied by a source peer
every time its neighbor list is updated or an updated neighbor cost 
table is received.
%%
%\paragraph*{Selective Flooding (SF)}
%
% TODO: I DON'T REMEMBER WHY THE FOLLOWING IS MENTIONING LTM WHICH IS ANOTHER
%       ALGORITHM. MAYBE THIS CAN BE USED AS A PART OF DISCUSSION AND/OR
%       COMPARISON OF AOTO AND LTM
% TODO: EDIT: LTM SEEMS A MISTAKE HERE BETTER NOT USE IT FOR DISCUSSION (THIS
%             IS AOTO)
%LTM's SF effectiveness has be proven to be detached from the different physical
% or overlay topologies. On the other hand, SF is more effective with large
%number of logical neighbors. It can reach an average optimization rate of 87.4
%percent on a logical topology with an average of 30 logical neighbors.
%
%\paragraph*{Active Topology (AT)}
%
%Different numbers of average logical neighbors has little to do with the
% effectiveness of AT. If the source has $n$ non-flooding peers, there are $n$
%potential neighbor replacements. The overhead to exhaust all $n$ possible
%replacements can be high, so in practice, after each replacement the source
%peer can decide whether it needs to find another candidate peer. This is done
%by computing the cost improvement ratio greater than some predefined
%termination threshold. The larger the threshold, the slower, in the number of
%optimization steps, the reduction of the normalized average distance. As a
%whole the average response time is significantly reduced when more optimization
%steps taken.
%%
\emph{AOTO}'s performance regarding the $3$ criteria is as follows:
%%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
medium &
% tracking of communication costs to all peer's neighbors as well as between
% any pair of neighbors. Whenever a new neighbor cost table is received or
% there is a change of neighbors, the source peer has to re-calculate the
% multicast tree and apply the randomized AT algorithm.
high &
% needs global knowledge
medium
\end{tabular}
}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The main objective of the \emph{Biased Neighbor Selection (BNS)} 
approach~\cite{BCCMSBZ2006} is to 
strengthen {\sl BitTorrent}'s~\cite{c_bittorrent_2003} locality 
by carefully choosing most of the neighbors of a peer to come from 
the same \emph{ISP}, while simultaneously adhering to the near-optimal
download performance of the protocol.
%%
By and large, {\sl BitTorrent} deploys mechanisms that have proved
aggressive to \emph{ISP} networking and accounting as they function 
in a ``without-borders'' approach. 
The default {\sl BitTorrent} specification
designates $35$ connections for each peer regardless if such connections are
placed within or outside the borders of the \emph{ISP} that hosts the node.
%%
\emph{BNS} proposes that only $k$ out of these $35$ connections be
established with nodes beyond the {\it ISP}-borders;
the $k$ connections offer an extended and ``global'' view of 
the network at large and seek to strike a balance between 
the load imposed inside and outside the \emph{ISP}.
The latter refrains peers from exchanging traffic through
expensive network links should there exist alternative local connections
offering the required services. 
%%
\emph{BNS} is realized through 
\begin{inparaenum}[1)]
\item
the modification of the tracker so that 
\emph{ISP}-local resources are rapidly identified
\item
the use of \p--traffic--shaping devices on \emph{ISP} edge--routers. 
\end{inparaenum}
%%
In~\cite{BCCMSBZ2006}, a combination of 
biased neighbor selection along with bandwidth throttling
and caching techniques for near-optimal lookup results is proposed. 
The effectiveness of \emph{BNS} is debated 
in~\cite{RTLCGZ2010} where experimental results suggest that 
most peers in a swarm reside in different autonomous systems (\emph{AS}s).
Consequently, clients frequently do not have enough peers to leverage
intra-\emph{AS} connections \cite{RTLCGZ2010}.
Moreover, \emph{BNS} deployment
requires knowledge of \emph{ISP} network mappings and awareness of possible
changes that occur in the infrastructure limiting its widespread adoption; the
following table qualitatively depicts the expected behavior of \emph{BNS}:
%%
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% inter-ISP cost minimized while retaining the near optimal performance of
% BitTorrent download.
high &
% this info is taken at swarm join
low &
% trackers should maintain and update proximity info/proximity info is
% retrieved from special hardware.
medium
\end{tabular}
}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Locality-Awareness in BitTorrent-like P2P Applications}
In~\cite{LCLX2009}, three techniques that 
``inject'' locality awareness in {\sl BitTorrent}-like applications are discussed.
The first \emph{macroscopic-level} technique
focuses on improving the neighbor selection process. 
When a peer asks a tracker to join, 
the later sorts its swarm peers according to
their hop-count distance from the requesting peer 
send out the top-$k$ list of nodes (e.g., $k$=$50$).
%%
The second technique, applied in an \emph{intermediate-level}, manipulates
{\sl BitTorrent}'s chocking/unchocking mechanism. A peer unchockes its $4$
closest, in terms of 
autonomous system hop-count, interested neighbors. 
The same applies also when the peer turns to 
the seeding state;  this  \emph{intermediate-level} approach favors
least distance and is in contrast to the original {\sl BitTorrent} 
implementation which favors uploading speed.
%%
The last technique operates at a \emph{microscopic-level} by choosing to pick
the next chunk to download based on a locality-first policy; this picks first
the chunk that is closer, as opposed to {\sl BitTorrent}'s rare-first chunk
picking policy.
%%VM I once more reworded the above, seemed to complex to me
The distance value of a chunk is computed as the mean
value of the distances of the peers that posses it.
Through experimentation on a real-world Internet topology simulated in
{\sf PlanetLab}, the locality-based approach achieves traffic optimization.
Nevertheless, it does not do that well when compared with the
random approach of the standard {\sl BitTorrent} protocol.

@@

In what follows, we outline how the surveyed protocols 
use elements of the above four methodologies. 
We then  offer a summary qualitative comparison for all surveyed
approaches applicable for \emph{unstructured}~\p~\emph{systems}.
Specifically, Table~\ref{unstructured:table} summarizes, for each
effort, the methodologies it incorporates, its special highlights as well as 
a rough estimation of the pros and cons in its implementation. 
Last but not least, Figures~\ref{unstructured:plot:efficiency}, \ref{unstructured:plot:overhead}
and \ref{unstructured:plot:scalability} pictorially compare the
surveyed techniques in terms of \emph{efficiency}, \emph{overhead} and \emph{scalability};
we defined these three criteria in Subsection~\ref{background:motivation}
and we collectively used them as a yardstick to qualitatively 
evaluate the presented approaches.

@@

In landmark-based algorithms, nodes use network delay (e.g., \emph{RTT}) as a
distance measurement method to position themselves with respect to ``a priori''
known servers on the Internet, like \emph{Ono} which uses the \emph{CDN}-infrastructure
for this purpose.
For example, in Figure~\ref{figure:landmarking}, the newly
arriving peer, measures its distance to an array of landmark points denoted as
$L_n$. Then it creates an ordered list of the landmarks based on its measured
distances uses that information to select its neighbors. Specifically,
the landmark--servers are used by nodes to estimate 
their positions based on the intuitive assumption that peers
with similar distances to a set of landmarks, are physically close to each
other, as well, over the network.  
%%
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.34]{img/pdf/landmarking.pdf}
\caption{Landmark binning during node bootstrap.}
\label{figure:landmarking}
\end{figure}
